---
title: "Asignment 5"
author: "Jay Oza"
date: "2022-11-28"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(cluster)
library(caret)
library(dendextend)
library(factoextra)
```

```{r}
c<-read.csv("/Users/jay/Downloads/Cereals.csv")
c<- na.omit(c)
head(c)
```

#Normalizing the dataset
```{r}
c<- c[4:16]
c <- scale(c,center = TRUE,scale = TRUE)
head(c)
```
#Task 1: Use Euclidean distance to apply hierarchical clustering to the data using the normalized measurements.Agnes can be used to compare the clustering from Ward, complete linkage, average linkage, and single linkage. Pick the best approach.
```{r}
Euclidean_Dist <- dist(c, method = "euclidean")
# Hierarchical clustering using Complete Linkage
hc1 <- hclust(Euclidean_Dist, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
round(hc1$height, 3)
```


# Compute with agnes and with different linkage methods
```{r}
hc_single <- agnes(c, method = "single")
print(hc_single$ac)
hc_complete <- agnes(c, method = "complete")
print(hc_complete$ac)
hc_average <- agnes(c, method = "average")
print(hc_average$ac)
hc_ward <- agnes(c, method = "ward")
print(hc_ward$ac)
```
#The agglomerative coefficient obtained by Ward's method is the largest. 

#visualizing  the dendrogram
```{r}
hc_Ward <- agnes(Euclidean_Dist, method = "ward")
pltree(hc_Ward, cex = 0.6, hang = -1, main = "Dendrogram of agnes for ward") 
```

#Task 2.How many clusters would you choose?
```{r}
#The largest difference in height can be used to determine the k value, hence K =5 is the best option.
hc_Ward <- hclust(Euclidean_Dist,method = "ward.D2")
clust_comp <- cutree(hc_Ward, k=5)
table(clust_comp)
plot(hc_Ward,cex=0.6)
rect.hclust(hc_Ward, k = 5, border = 2:10,)
Temp <- cbind(as.data.frame(cbind(c,clust_comp)))
#Visualizing the clusters in Scatter plot
fviz_cluster(list(data=Euclidean_Dist, cluster = clust_comp))
```

#Task 3: Choose a set of cereals for the elementary public schools to stock in their daily cafeterias. There is a different cereal available every day, but they should all promote a healthy diet. To complete this task, you must locate a group of "healthy cereals."
```{r}
Healthy_cereal <- na.omit(read.csv("/Users/Jay/Downloads/Cereals.csv"))
Healthy_cereal<- cbind(Healthy_cereal,clust_comp)
mean(Healthy_cereal[Healthy_cereal$clust_comp==1,"rating"])
mean(Healthy_cereal[Healthy_cereal$clust_comp==2,"rating"])
mean(Healthy_cereal[Healthy_cereal$clust_comp==3,"rating"])
mean(Healthy_cereal[Healthy_cereal$clust_comp==4,"rating"])
mean(Healthy_cereal[Healthy_cereal$clust_comp==5,"rating"])
```

It is clear that Cluster1 has maximum rating (73.84446), therefore we'll choose it as a healthy cereal.
